---
output: 
  html_document: 
    keep_md: yes
---
# Reproducible Research: Week 2 Assignment
by AC

#### Call the library
```{r echo=TRUE}
library(knitr)
```

#### Download the file
```{r echo=TRUE}
fil <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fil, destfile="./activity.zip")
```

#### Unzip the file
```{r echo=TRUE}
unzip(zipfile="./activity.zip")
```

## Read the downloaded CSV file
```{r echo=TRUE}
act <- read.csv("activity.csv")
```

#### Dimensions
```{r echo=TRUE}
dim(act)
```

#### Variables present
```{r echo=TRUE}
names(act)
```

#### Check classes of variables
```{r echo=TRUE}
str(act)
```

## Mean total steps per day (ignoring NAs)
```{r echo=TRUE}
library(dplyr)
perday <- group_by(act, date)
spd <- summarise(perday, tot=sum(steps))
```

### Histogram
```{r}
hist(spd$tot, col = "Blue", xlab = "Average no. of steps", ylab = "No. of days", main = "Total number of steps taken each day")
```

## Mean & Median
```{r echo=TRUE}
paste("Mean of steps per day: ",mean(spd$tot, na.rm = T))
paste("Median of steps per day: ",median(spd$tot, na.rm = T))
```

#### Create table containing average no. of steps per interval
```{r echo=TRUE}
gb <- group_by(act, interval)
avgint <- summarise(gb, avg=mean(steps, na.rm = T))
```

## Time Series plot
```{r echo=TRUE}
plot(avgint$interval, avgint$avg, type="l", xlab = "5 minute Intervals", ylab = "Average no. of steps", main = "Average no. of steps in each interval")
```

## Maximum avg steps across intervals
```{r echo=TRUE}
paste("Interval with maximum average steps per day: ", avgint$interval[avgint$avg %in% max(avgint$avg)])
```

### Imputing missing values
```{r echo=TRUE}
paste("No. of NAs in data set: ", nrow(act[act$steps %in% NA,]))
```

## Strategy to replace NA values: Replace with Mean for that Interval
```{r echo=TRUE}
for(i in 1:nrow(act)){
    if(act$steps[i] %in% NA)
    {
        act$steps[i]=avgint$avg[act$interval[i] == avgint$interval]
    }
}
paste("No. of NAs after update: ", nrow(act[act$steps %in% NA,]))
```

Dimensions of updated data set
```{r echo=TRUE}
dim(act)
```

## Histogram for total no. of steps per day
```{r echo=TRUE}
perday <- group_by(act, date)
nona <- summarise(perday, total=sum(steps))
hist(nona$total, col = "Red", xlab = "Average no. of steps", ylab = "No. of days", main = "Total number of steps taken each day: No NAs")
```

## Mean
```{r}
paste("Mean no. of steps per day: ", mean(nona$total))
```

## Median
```{r}
paste("Median no. of steps per day: ", median(nona$total))
```

There is no significant change in values of mean and median despite the imputation.


#### Change date from factor to date data type
```{r echo=TRUE}
act$date <- as.Date(act$date)
class(act$date)
day <- vector()
for(i in 1:nrow(act)){
    if(weekdays(act$date[i]) %in% "Monday"|weekdays(act$date[i]) %in% "Tuesday"|weekdays(act$date[i]) %in% "Wednesday"|weekdays(act$date[i]) %in% "Thurday"|weekdays(act$date[i]) %in% "Friday")
    {
        day[i] <- "Weekday"
    }
    else
    {
        day[i] <- "Weekend"
    }
}
act <- cbind(act,day)
act$day <- as.factor(act$day)
levels(act$day)
```

#### Separating weekdays and weekend data
```{r echo=TRUE}
wd <- act[act$day %in% "Weekday",]
we <- act[act$day %in% "Weekend",]
gbwd <- group_by(wd, interval)
gbwe <- group_by(we, interval)
avgwd <- summarise(gbwd, avg=mean(steps))
avgwe <- summarise(gbwe, avg=mean(steps))
```

## Panel Time Series Plots: Weekdays & Weekend
```{r echo=TRUE}
par(mfrow=c(2,1))
plot(avgwd$interval, avgwd$avg, type="l", xlab = "5 minute Intervals", ylab = "Average no. of steps", main = "Average no. of steps in each interval: Weekdays", ylim = c(0,250))
plot(avgwe$interval, avgwe$avg, type="l", xlab = "5 minute Intervals", ylab = "Average no. of steps", main = "Average no. of steps in each interval: Weekend", ylim = c(0,250))
```



